---
title: "DATA606 - Final Meetup"
author: Jason Bryer, Ph.D.
date: December 8, 2017
knit: (function(inputFile, encoding) { input.dir <- normalizePath(dirname(inputFile)); rmarkdown::render(input = inputFile, encoding = encoding, quiet=FALSE, output_file = paste0(input.dir,'/../docs/slides/', tools::file_path_sans_ext(basename(inputFile)), '.html')); })
output:
  ioslides_presentation:
    self_contained: true
    widescreen: true
    smaller: true
---
	
<div class="notes">
Documentation on using ioslides is available here:
http://rmarkdown.rstudio.com/ioslides_presentation_format.html
Some slides are adopted (or copied) from OpenIntro: https://www.openintro.org/
</div>

```{r setup, echo=FALSE, results='hide', warning=FALSE, message=FALSE}
set.seed(2112)
library(ggplot2)
library(openintro)
library(IS606)
library(reshape2)
library(psych)
library(reshape2)
library(xtable)
options(width=100)
par(mar=c(2.5,1,2,1))

```

## Final Exam

* Will be available tomorrow.

* Due by end of day December 15th.

* You may use your book and course materials.

* There are two parts:
	1. Part one has some conceptual questions.
	2. Part two has a small data set to analyze with R, then answer some interpretation questions.

## Presentations

* [Bin Lin](http://rpubs.com/blin261/233806) (7.39)
* [Jenny Xie](http://rpubs.com/Jenny/Data606_Meetup) (8.3)
* [Todd Weigel](http://rpubs.com/tmuse987/233757) (7.41)
* [Liam Byrne](http://rpubs.com/Liam-O/Data606_Presentation) (7.37)
* [Ahsanul Choudhury](http://rpubs.com/choudhury1023/232893) (6.19)


## The Logistic Function {.flexbox .vcenter}

$$ \sigma \left( t \right) =\frac { { e }^{ t } }{ { e }^{ t }+1 } =\frac { 1 }{ 1+{ e }^{ -t } }  $$

```{r, fig.width=6, fig.height=3}
logistic <- function(t) { return(1 / (1 + exp(-t))) }
df <- data.frame(x=seq(-4, 4, by=0.01))
df$sigma_t <- logistic(df$x)
plot(df$x, df$sigma_t)
```

## *t* as a Linear Function

$$ t = \beta_0 + \beta_1 x $$

The logistic function can now be rewritten as

$$ F\left( x \right) =\frac { 1 }{ 1+{ e }^{ -\left( { \beta  }_{ 0 }+\beta _{ 1 }x \right)  } } $$

We use the same ordinary least squares procedure we used before. That is, we wish to minimize the sum of squared residuals.

## Example: Hours Studying Predicting Passing

```{r}
study <- data.frame(
	Hours=c(0.50,0.75,1.00,1.25,1.50,1.75,1.75,2.00,2.25,2.50,2.75,3.00,3.25,3.50,4.00,4.25,4.50,4.75,5.00,5.50),
	Pass=c(0,0,0,0,0,0,1,0,1,0,1,0,1,0,1,1,1,1,1,1)
)
lr.out <- glm(Pass ~ Hours, data=study, family=binomial)
lr.out
```

## Fitted Values

```{r}
study[1,]
logistic <- function(x, b0, b1) {
	return(1 / (1 + exp(-1 * (b0 + b1 * x)) ))
}
logistic(.5, b0=-4.078, b1=1.505)
```

Of course, the `fitted` function will do the same:

```{r}
study$fitted <- fitted(lr.out)
study[1,]
```

## Plotting the Results {.flexbox .vcenter}

```{r, fig.height=4}
binomial_smooth <- function(...) {
	geom_smooth(method = "glm", method.args = list(family = "binomial"), ...)
}

ggplot(study, aes(x=Hours, y=Pass)) + geom_point() + 
	binomial_smooth(se=FALSE)
```

## Residuals {.flexbox .vcenter}

```{r, fig.height=4}
study$residuals <- study$Pass - study$fitted

ggplot(study, aes(x=Hours, y=Pass)) + geom_point() + 
	binomial_smooth(se=FALSE) +
	geom_segment(aes(xend=Hours, yend=fitted), color='maroon')
```


## My Work

My research interest is in propensity score methods. Propensity score analysis (PSA) is a quasi-experimental design used to estimate causality from observational studies. It is generally conducted in two phases:

1. Estimate propensity scores (i.e. probability of being in the treatment) using the observed covariates.
	a. Check balance
	b. Re-estimate propensity scores
2. Estimate effect sizes using typical group differences (e.g. t-tests)

Areas I have worked on:

* Multilevel PSA (see [`multilevelPSA`](http://jason.bryer.org/multilevelPSA) R package)
* Matching with non-binary treatments (see [`TriMatch`](http://jason.bryer.org/TriMatch) R package)
* Bootstrapping PSA (see [`PSAboot`](http://jason.bryer.org/PSAboot) R package)

## Thank You

This has been a great semester. Please don't hesitate to reach out:

* Email: jason.bryer@gmail.com
* Github: http://github.com/jbryer
* Blog: http://bryer.org
* [LinkedIn](https://www.linkedin.com/profile/view?id=AAMAAATGdnoBOWXg80yqna6fSkgnZdabZP7Ck9w&trk=hp-identity-name)
* Twitter: [@jbryer](https://twitter.com/jbryer) (don't Tweet very much though)


