---
title: "DATA606 - Foundation for Inference"
author: Jason Bryer, Ph.D.
date: October 6, 2016
knit: (function(inputFile, encoding) { input.dir <- normalizePath(dirname(inputFile)); rmarkdown::render(input = inputFile, encoding = encoding, quiet=FALSE, output_file = paste0(input.dir,'/../docs/slides/', tools::file_path_sans_ext(basename(inputFile)), '.html')); })
output:
  ioslides_presentation:
    self_contained: true
    widescreen: true
    smaller: true
---

<div class="notes">
Documentation on using ioslides is available here:
http://rmarkdown.rstudio.com/ioslides_presentation_format.html
Some slides are adopted (or copied) from OpenIntro: https://www.openintro.org/
</div>

```{r setup, echo=FALSE, results='hide', warning=FALSE, message=FALSE}
set.seed(2112)
library(ggplot2)
library(openintro)
library(IS606)
par(mar=c(2.5,1,2,1))

PlotDist <- function(alpha, from = -5, to = 5, n = 1000, filename = NULL,
    alternative = c("two.tailed", "greater", "lesser"), 
    distribution = c("normal", "t", "F", "chisq", "binomial"), 
    colour = "black", fill = "skyblue2",
    ...)
{
    alternative <- match.arg(alternative)
    alt.alpha <- switch(alternative, two.tailed = alpha/2, greater = alpha,
        lesser = alpha)
    MyDen <- switch(distribution, normal = dnorm, t = dt, F = df,
        chisq = dchisq, binomial = dbinom)
    MyDist <- switch(distribution, normal = qnorm, t = qt, F = qf,
        chisq = qchisq, binomial = qbinom)
    crit.lower <- MyDist(p = alt.alpha, lower.tail = TRUE, ...)
    crit.upper <- MyDist(p = alt.alpha, lower.tail = FALSE, ...)
    cord.x1 <- c(from, seq(from = from, to = crit.lower, length.out = 100),
        crit.lower)
    cord.y1 <- c(0, MyDen(x = seq(from = from, to = crit.lower,
        length.out = 100), ...), 0)
    cord.x2 <- c(crit.upper, seq(from = crit.upper, to = to,
        length.out = 100), to)
    cord.y2 <- c(0, MyDen(x = seq(from = crit.upper, to = to,
        length.out = 100), ...), 0)
    if (!is.null(filename)) pdf(file = filename)
    curve(MyDen(x, ...), from = from, to = to, n = n, col = colour,
        lty = 1, lwd = 2, ylab = "Density", xlab = "Values")
    if (!identical(alternative, "greater")) {
        polygon(x = cord.x1, y = cord.y1, col = fill)
    }
    if (!identical(alternative, "lesser")) {
        polygon(x = cord.x2, y = cord.y2, col = fill)
    }
    if (!is.null(filename)) dev.off()
}
```

## Announcements

* No meetup next week or November 3rd (for those signed up for Nov 3, I moved you to Nov 10).
* Labs 0, 1, and 2 have been graded.
* Homework 1, 2, and 3 have been graded.


## Meetup Presentations



## Type I and II Errors

* Type I Error: **Rejecting** the null hypothesis when it is **true**.
* Type II Error: **Failing to reject** the null hypothesis when it is **false**.


## Null Distribution {.flexbox .vcenter}

```{r, fig.width=6, fig.height=3.5}
(cv <- qnorm(0.05, mean=0, sd=1, lower.tail=FALSE))
PlotDist(alpha=0.05, distribution='normal', alternative='greater')
abline(v=cv, col='blue')
```

## Alternative Distribution {.flexbox .vcenter}

```{r, fig.width=6, fig.height=3.5}
cord.x1 <- c(-5, seq(from = -5, to = cv, length.out = 100), cv)
cord.y1 <- c(0, dnorm(mean=cv, x=seq(from=-5, to=cv, length.out = 100)), 0)
curve(dnorm(x, mean=cv), from = -5, to = 5, n = 1000, col = "black",
        lty = 1, lwd = 2, ylab = "Density", xlab = "Values")
polygon(x = cord.x1, y = cord.y1, col = 'lightgreen')
abline(v=cv, col='blue')
```

```{r}
pnorm(cv, mean=cv, lower.tail = FALSE)
```

## Another Example (mu = 2.5) {.flexbox .vcenter}

```{r}
mu <- 2.5
(cv <- qnorm(0.05, mean=0, sd=1, lower.tail=FALSE))
```

```{r, echo=FALSE, fig.width=3.5, fig.height=3.5, fig.show='hold'}
pv <- pnorm(mu, mean=0, sd=1, lower.tail=FALSE)

PlotDist(alpha=pv, distribution='normal', alternative='greater')
abline(v=mu, col='blue')
title('Null Distribution')

cord.x1 <- c(-5, seq(from = -5, to = cv, length.out = 100), cv)
cord.y1 <- c(0, dnorm(mean=mu, x=seq(from=-5, to=cv, length.out = 100)), 0)
curve(dnorm(x, mean=mu), from = -5, to = 5, n = 1000, col = "black",
        lty = 1, lwd = 2, ylab = "Density", xlab = "Values")
polygon(x = cord.x1, y = cord.y1, col='lightgreen')
abline(v=mu, col='blue')
title('Alternative Distribution')
```

## Numeric Values

Type I Error

```{r}
pnorm(mu, mean=0, sd=1, lower.tail=FALSE)
```

Type II Error

```{r}
pnorm(cv, mean=mu, lower.tail = TRUE)
```

## Shiny Application

Visualizing Type I and Type II errors: [http://shiny.albany.edu/stat/betaprob/](http://shiny.albany.edu/stat/betaprob/)


